---
title: "PEVS and Energy Security: Systematic Review"
author: "Rocio Uria-Martinez"
date: "07/10/2023"
output: html_document
---

```{r LoadLibraries, eval = T}

library(tidyverse)
library(readxl)
library(scales) 
library(purrr)
```

```{r PlotFormatTheme, eval = T}

EERE.palette.long <- c("#007934", "#5E6A71", "#E37222", "#FECB00", "#69BE28", "#1F82BB", "#005B82", "#C3C8C8", "#9865AA", "#9E2432")
EERE.palette.names <- c("DarkGreen", "DarkGrey", "DarkOrange", "EggYolk", "PistachioGreen", "AquaBlue", "DarkBlue", "LightGrey", "Purple", "DarkRed")
############################################################################

##custom plot theme
myFGtheme <- theme_bw() +   
  theme(text=element_text(family="Franklin Gothic Book", size=12)) +
  theme(axis.text = element_text(size=9), axis.title=element_text(size=9, face="bold", family = "Franklin Gothic Medium")) +
  #theme(axis.text.x = element_text(angle = 65, vjust = .95,hjust=.95))+
  theme(panel.grid.major.x = element_blank(), panel.grid.minor.x = element_blank(),
        panel.grid.major.y = element_line(color = "grey", linewidth = 0.5), panel.grid.minor.y = element_line(color = "white"),
        panel.background = element_blank(),
        strip.background = element_rect(color = "white"),
        axis.line = element_line(colour = "grey", linewidth = 0.5), panel.border = element_blank()) +
  theme(legend.direction = "horizontal", legend.position = "bottom", legend.box = "vertical", legend.title.align = 0, legend.key=element_blank()) + 
  theme(legend.title=element_text(size = 12, family = "Franklin Gothic Medium", face = "bold"))
```

Searches in Scopus:
https://www.scopus.com/search/form.uri?display=basic#basic

Searches in Web of Science:
https://www.webofscience.com/wos/woscc/advanced-search

The initial queries looked for records published from 2010 on that included in the title, keywords, or abstract "PEV" or "PEVs" or "PHEV" or "PHEVs" or "EV" or "EVs" or "electric vehicle" or "electric vehicles" and "energy security" 

Note: the data with a _2 suffix comes from later expanded search for "BEV" or "BEVs" or "ZEV" or "ZEVs" and "energy security". Google Scholar returned zero records for that expanded search

```{r readData, eval = T}

scopus <- read.csv("../Data/systematic_review/scopus_PEV_ES_20230710.csv") %>%
  mutate(in_scopus = 1)

#387 entries of which 44 do not have DOI

scopus_2 <- read.csv("../Data/systematic_review/scopus_BEVZEV_ES_20230724.csv") %>%
  mutate(in_scopus = 1)
#26 entries of which 2 do not have DOI


webofscience <- read_excel("../Data/systematic_review/savedrecs_WoS_PEV_ES_20230710.xls") %>%
  mutate(in_wos = 1)

#247 entries out of which 47 do not have DOI

webofscience_2 <- read_excel("../Data/systematic_review/savedrecs_WoS_BEVZEV_ES_20230724.xls") %>%
  mutate(in_wos = 1)
#13 entries of which 0 do not have DOI

googlescholar <- read.csv("../Data/systematic_review/GoogleScholarcitations_PEV_ES_20230710.csv") %>%
  mutate(in_scholar = 1)

#Google Scholar citations do not include DOI

```

```{r selectUnique, eval = T}
scopus_select <- scopus %>%
  select(Authors, Title, Year, Source.title, Volume, Issue, Page.start, Page.end, DOI, Abstract, in_scopus) %>%
  rename(SourceTitle = Source.title, StartPage = Page.start, EndPage = Page.end)

scopus_select_DOI <- scopus_select %>%
  filter(DOI != "") %>%
  mutate_at(c("Volume", "Issue", "StartPage", "EndPage"), as.integer) %>%
  filter(!(Authors == "Cornils H." & is.na(Volume))) %>%
  filter(!(Authors == "Veenstra M.J.; Gearhart C." & is.na(Volume))) %>%
  filter(!(Authors == "Eberle U.; Von Helmolt R." & SourceTitle == "Electric and Hybrid Vehicles: Power Sources, Models, Sustainability, Infrastructure and the Market"))

scopus_select_noDOI <- scopus_select %>%
  filter(DOI == "") %>%
  mutate_at(c("Volume", "Issue", "StartPage", "EndPage"), as.integer) %>%
  mutate(database = "scopus") %>%
  select(-in_scopus)

webofscience_select <- webofscience %>%
  select(Authors, `Article Title`, `Publication Year`,`Source Title`, Volume, Issue, `Start Page`, `End Page`, DOI, Abstract, in_wos) %>%
  rename(Title = `Article Title`, Year = `Publication Year`, SourceTitle = `Source Title`, StartPage = `Start Page`, EndPage = `End Page`)

webofscience_select_DOI <- webofscience_select %>%
  filter(DOI != "") %>%
  mutate_at(c("Issue", "EndPage"), as.integer)

webofscience_select_noDOI <- webofscience_select %>%
  filter(is.na(DOI)) %>%
  mutate_at(c("Issue", "EndPage"), as.integer) %>%
  mutate(database = "WoS") %>%
  select(-in_wos)

googlescholar_select <- googlescholar %>%
  select(Authors, Title, Publication, Volume, Number, Pages, Year) %>%
  separate(Pages, into = c("StartPage", "EndPage"), sep = "-") %>%
  rename(SourceTitle = Publication, 
         Issue = Number) %>%
  mutate(Abstract = NA,
         DOI = NA,
         database = "scholar")

###Remove duplicate entries by DOI

scopus_wos_DOI_join <- scopus_select_DOI %>%
  full_join(webofscience_select_DOI[,c("DOI", "in_wos")], by = "DOI") 

unique_DOIs_notinscopus <- scopus_wos_DOI_join %>%
  filter(is.na(in_scopus)) %>%
  pull(DOI)

webofscience_select_DOI_notinscopus <- webofscience_select_DOI %>%
  filter(DOI %in% unique_DOIs_notinscopus)

unique_DOIs <- rbind(scopus_select_DOI[-c(11)], webofscience_select_DOI_notinscopus[-c(11)]) 

#This should be an empty set if I successfully removed the duplicates
dup_DOIs <- unique_DOIs %>%
  group_by(DOI) %>%
  summarise(count = n()) %>%
  filter(count == 2) %>%
  pull(DOI)

#This should have zero rows if I successfully removed the duplicates
view_dup_records <- unique_DOIs %>%
  filter(DOI %in% dup_DOIs)

###Remove duplicate entries based on other attributes (for entries without DOIs)

scopus_wos_scholar_nonDOI_join <- rbind(scopus_select_noDOI, webofscience_select_noDOI, googlescholar_select) %>%
  filter(Authors != "") %>%
  separate(Authors, into = c("firstauthor", "rest"), sep = " ", remove = FALSE) %>%
  select(-rest) %>%
  mutate(firstauthor = gsub(",", "", firstauthor),
         label = paste(firstauthor, Year, sep = "_")) %>%
  filter(!(firstauthor == "Horn" & database == "WoS")) %>%
  filter(!(label == "Kang_2015" & database == "WoS")) %>%
  filter(!(firstauthor == "Popczyk" & database == "WoS")) %>%
  filter(!(firstauthor == "Shiiki"  & Volume == 2)) %>%
  filter(!(label == "Schmidt_2013" & is.na(Volume))) %>%
  filter(!(label == "Sovacool_2010" & is.na(Volume))) %>%
  filter(!(label == "Zhang_2015" & EndPage == "Jan"))

#This has one remaining row from one instance in which two papers have the same author_year label 
dup_labels <- scopus_wos_scholar_nonDOI_join %>%
  group_by(label) %>%
  summarise(count = n()) %>%
  filter(count > 1) %>%
  pull(label)

#This has two remaining rows from one instance in which two papers have the same author_year label 
view_dup_records <- scopus_wos_scholar_nonDOI_join %>%
  filter(label %in% dup_labels)

#scopus_wos_DOI_join_2 <- scopus_wos_DOI_join %>%
#  select(-in_scopus, -in_wos)

scopus_wos_scholar_nonDOI_join_2 <- scopus_wos_scholar_nonDOI_join %>%
  select(-label, -firstauthor, -database)

scopus_wos_scholar_unique <- rbind(unique_DOIs, scopus_wos_scholar_nonDOI_join_2) %>%
  arrange(Year, Authors)

write.csv(scopus_wos_scholar_unique, "../Output/scopus_wos_scholar_unique.csv", row.names = F)

```

```{r selectUnique_BEVZEV, eval = T}
scopus_2_select <- scopus_2 %>%
  select(Authors, Title, Year, Source.title, Volume, Issue, Page.start, Page.end, DOI, Abstract, in_scopus) %>%
  rename(SourceTitle = Source.title, StartPage = Page.start, EndPage = Page.end)

scopus_2_select_DOI <- scopus_2_select %>%
  filter(DOI != "") %>%
  mutate_at(c("Volume", "Issue", "StartPage", "EndPage"), as.integer) 

scopus_2_select_noDOI <- scopus_2_select %>%
  filter(DOI == "") %>%
  mutate_at(c("Volume", "Issue", "StartPage", "EndPage"), as.integer) %>%
  mutate(database = "scopus") %>%
  select(-in_scopus)

webofscience_2_select <- webofscience_2 %>%
  select(Authors, `Article Title`, `Publication Year`,`Source Title`, Volume, Issue, `Start Page`, `End Page`, DOI, Abstract, in_wos) %>%
  rename(Title = `Article Title`, Year = `Publication Year`, SourceTitle = `Source Title`, StartPage = `Start Page`, EndPage = `End Page`)

webofscience_2_select_DOI <- webofscience_2_select %>%
  filter(DOI != "") %>%
  mutate_at(c("Issue", "EndPage"), as.integer)

webofscience_2_select_noDOI <- webofscience_2_select %>%
  filter(is.na(DOI)) %>%
  mutate_at(c("Issue", "EndPage"), as.integer) %>%
  mutate(database = "WoS") %>%
  select(-in_wos)

###Remove duplicate entries by DOI

scopus_wos_2_DOI_join <- scopus_2_select_DOI %>%
  full_join(webofscience_2_select_DOI[,c("DOI", "in_wos")], by = "DOI") 

unique_DOIs_notinscopus_2 <- scopus_wos_2_DOI_join %>%
  filter(is.na(in_scopus)) %>%
  pull(DOI)

webofscience_select_DOI_notinscopus_2 <- webofscience_2_select_DOI %>%
  filter(DOI %in% unique_DOIs_notinscopus_2)

unique_DOIs_2 <- rbind(scopus_2_select_DOI[-c(11)], webofscience_select_DOI_notinscopus_2[-c(11)]) 

#This should be an empty set if I successfully removed the duplicates
dup_DOIs_2 <- unique_DOIs_2 %>%
  group_by(DOI) %>%
  summarise(count = n()) %>%
  filter(count == 2) %>%
  pull(DOI)

#This should have zero rows if I successfully removed the duplicates
view_dup_records_2 <- unique_DOIs_2 %>%
  filter(DOI %in% dup_DOIs_2)

scopus_wos_2_unique <- rbind(unique_DOIs_2, scopus_2_select_noDOI[-c(11)]) %>%
  arrange(Year, Authors)

write.csv(scopus_wos_2_unique, "../Output/scopus_wos_2_unique.csv", row.names = F)

```

None of the references from the expanded search (BEV/ZEV) is unique relative to the initial queries and relevant to our analysis of energy security implications of PEVs.

```{r TopicAnalysis, eval = T}
screenedrefs <- read.csv("../Output/scopus_wos_scholar_unique_rev.csv") %>%
  filter(Keep_prescreen == 1) %>%
  group_by(Topic) %>%
  summarise(count = n())

```

